---
title: 'Final Report: Selling to both Sides (Moore 2012)'
author: "Carlo Greß"
date: "26 5 2022"
output:
  pdf_document: default
  html_document: default
bibliography: references.bib 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
library(tidyverse)
library(tibble)
library(ggdag)
```

## Loading the data 

The replication data set for Matthew Moore's article "Selling to both sides" is available [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/8PKAI0)

The original article is available [here](https://www.tandfonline.com/doi/pdf/10.1080/03050629.2012.676511?casa_token=JNzRApl8w24AAAAA:Sth2YQPRDCYC-RSAX90DJVyzofnIauIWa9eU7561mATdM_fm2NdfEJHqKdOOavQeNwlJNLF-mvJExw ) 

```{r}
load("Replication files\\moore selling to both sides Internaitonal Interactions.RData")

moore_data <- x
```


# Introduction / Summary Moore 
This final project provides a methodological critique of the article "Selling to both sides: The Effects of Major Conventional Weapons Transfers on Civil War Severity and Duration" [@moore2012selling]. After a brief summary of the research objective and hypothesis, the used data, and results, I will discuss Moore examines how transfers of certain types of weapons impact both a civil war's duration as well as its severity. More specifically, the article distinguishes between weapon transfers to rebel groups and weapon transfers to governments and compares potentially deviating effects. Regarding weapon characteristics, only a certain weapon type is considered: Major conventional weapons (MCW). Since the author uses the Stockholm International Peace Research Institute's (SIPRI) definition of MCW, this most importantly includes aircraft, armoured vehicles, artillery, sensors, air defense systems, missiles, and ships. On the other hand, small arms as pistols and machine guns are excluded from the analysis. 
Using data from the Peace Research Institute Oslo (PRIO) on conflicts and SIPRI data on weapon transfers, Moore examines the effect of transfers of these weapons to both rebels and governments on two distinct dependent variables: the conflict's severity and the conflict's duration. Severity is measured by the number of battle-related deaths during the respective conflict. Duration is measured in whole years. He tests four hypotheses: Regarding the severity of civil conflicts, he states that MCW transfers to both rebel groups (H1) as well as to governments (H2) will increase the severity of a conflict (thus increase the number of battle-related deaths). These hypotheses arise from the expected causal mechanism that MCW have the potential to escalate a civil conflict, especially when rebels are able to acquire them. On the other hand, and although Moore acknowledges a potentially deterring effect of MCW in the hands of governments, he argues that MCW contribute to a higher severity once a conflict broke out. Ultimately, he concludes that both transfers to rebels and to governments will increase the severity of a civil conflict. 
Regarding the duration of civil conflicts, Moore expects diverging effects of MCW transfers when rebels and governments are considered: MCW transfers should decrease a conflict’s duration (H3), while transfers to governments are expected to increasing the duration (H4). Concerning the expected conflict-shortening effect of MCW transfers to rebels, the author argues that MCW increase the rebels’ fighting capacity, which ultimately results in a more credible threat to governments. Subsequently, governments may be willing to negotiate peace settlements, which would ultimately terminate the conflict and shorten conflict duration. The last hypothesis is underpinned by the expectation that additional weapons are only marginally improving a government’s relative strength.

# Applied method
In order to test these four hypotheses, Moore specifies several models. For the hypotheses on conflict severity, the author ran five distinct ordinary least squares regression models with diverging specifications of the weapon import variable. In order to test the hypotheses on conflict duration, Moore specified a Cox proportional hazard model. In both the OLS regression as well as in the cox regression, several control variables that have been found significant in previous literature examining similar problems are included. Moore controls for population size, military quality, GDP, a binary cold war indicator, a binary indicator indicating mountainous terrain, a binary democracy indicator, ethnic polarization, religious polarization, and a binary variable indicating whether there have been third-party interventions. Figure 1 shows an illustration of the expected causal mechanism in a directed acyclic graph (DAG). 

```{r}
dag = dagify(S ~ WT,
             S ~ CV,
             WT ~ CV)
ggdag(dag) +  theme_dag_blank()
```

# Method Review 
-	OLS regression
-	Cox regression
-	(Weibull regression)
-	Problems: Endogeneity
-	Coding mistake

```{r}
moore_replication$new_lstarmimp5 <- log(moore_replication$starmimp5)
```

```{r}
moore_replication_subset = moore_replication %>% 
  subset(sipricomp == 1) %>% 
  subset(conflict_name != "Vietnam War")
```

```{r}
model3 <- lm(lnbdb ~ lstarmimp5 + lrarmimp5 + lnduration + lnpop + lnmilqual + lngdp + cw + lnmountain + democ  + ethnicpolar + relpolar, data = moore_replication_subset)
```

```{r}
summary(model3)
```
```{r}
model5 <- lm(lnbdb  ~ lstarmimp5 + lrarmimp5 + lnduration + lnpop + lnmilqual + lngdp + cw + lnmountain + democ + ethnicpolar + relpolar + intervention, data = moore_replication_subset)

summary(model5)
```



# Improvement 
-	DAGs
-	Alternative (exponentiated) import variable 
-	Endogeneity solution (instrument)

# Conclusion 
